---
title: "writing functions"
author: "Shihui Peng"
date: "2023-11-14"
output: github_document
---
```{r, echo=FALSE, message=FALSE}
library(tidyverse)
library(rvest)
```

set seed for reproducibility
```{r}
set.seed(12345)
```
* we will generate random number today, this code help us get exactly the same random number every single time. this code tell it where to start.
  * to make sure random number generation starts at the same point and give us the same value.

# Z score function
z score subtract the mean and divide by the sd.
generate a vec of random samples:
```{r}
x_vec = rnorm(20, mean = 5, sd = .3)
```
* take a sample from a normal distribution. do 20 of this and set mean as 5 and sd as 0.3

compute z score for x_vec
```{r}
(x_vec - mean(x_vec)) / sd(x_vec)
```

suppose i have many vectors and want to quickly compute z scores for them.
* write a function to do this!
```{r}
z_score = function(x) {
# check these
    if (!is.numeric(x)) {
    stop("Argument x should be numeric")
  } else if (length(x) < 2) {
    stop("you need at least 2 numbers to get z scores")
  }

# if neither of above are true, then go as below:  
  z = (x - mean(x) )/ sd(x)
  z
}
```
* If I want to repeat this process for lots of samples, I need a function that takes the sample as an argument, computes the vector of Z scores in the body, and returns the result. I define such a function below.
  * name what that input is -- call it `x` inside `function()` (generic vector).
  * use that to produce z -- `z = (x - mean(x) / sd(x))`
  * return z -- `z`

check if this works:
```{r, error=TRUE}
z_score(x = x_vec)

z_score(x = rnorm(10, mean = 5))

# return NA, bc no sd when only 1 number (w/o the 'if' part in our function)
z_score(x = 3)

# return error bc no sd or mean for character variables (w/o the 'if' part in our function)
z_score(c('my', 'name', 'is', 'sheryl'))

# return 0.5  0.5 -1.5  0.5 bc r converts these to 1 and 0 (w/o the 'if' part in our function)
z_score(c(TRUE, TRUE, FALSE, TRUE))

# similar as above (w/o the 'if' part in our function)
z_score(sample(c(TRUE, FALSE), 25, replace = TRUE))

# return error bc cannot get mean and sd for the entire data frame (w/o the 'if' part in our function)
z_score(iris)
```




